---
title: "Ordinal regression analysis"
author: "Sergi Pujol Rigol"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readxl)
library(ggplot2)
library(brms)
library(VGAM)
library(ordinal)
library(MASS)
library(ordinalgmifs)
library(MuMIn)
```

#Dataset loading and processing

Load the Arthritis dataset from the multgee package

```{r}
library(multgee)
data(arthritis)
```

Process the dataset and select only the T=5 months follow-up time.

```{r}
#Baseline as factor
arthritis <- subset(arthritis, arthritis$time == 5) %>%
  mutate(across(c(y,sex,trt,baseline), as.factor)) %>%
  dplyr::select(-c(id,time)) %>%
  rename(bsl = baseline)

#Baseline as numeric
arthritis <- subset(arthritis, arthritis$time == 5) %>%
  mutate(across(c(y,sex,trt), as.factor)) %>%
  dplyr::select(-c(id,time)) %>%
  rename(bsl = baseline)

#Define the order of categories
ordre_y <- c("1", "2", "3", "4", "5")
arthritis$y <- ordered(arthritis$y, levels = ordre_y) #Mutate response y to ordered factor

#NA?
sum(is.na(arthritis))
```

Eliminate rows with missing values (NAs) from the dataset.

```{r}
arthritis <- arthritis[complete.cases(arthritis),]
```

#Dataset description

```{r}
str(arthritis)
summary(arthritis)
```

Bivariate correlation analysis with the ggpairs function from the GGally package.

```{r}
library(GGally)

ggpairs <- GGally::ggpairs(arthritis[,2:5])
ggpairs + theme(
  text = element_text(size = 20),
  axis.text.x = element_text(size = 8),
  axis.text.y = element_text(size = 8)
)
```

Descriptive bivariate plots

```{r}
#y vs treatment
ggplot(arthritis, aes(x = trt, y = y)) +
  geom_boxplot(size = .75) +
  geom_jitter(alpha = .5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

#y vs baseline for both treatments
dp <- ggplot(arthritis, aes(x=baseline, y=y, fill=trt)) + 
  geom_violin(trim=FALSE)+
  geom_boxplot(width=0.1, fill="white")+
  labs(x="Baseline", y = "Self-assessment arthritis")
dp + theme_classic() +
  coord_flip() +
  scale_fill_discrete(name = "Treatment")

#y vs age
ggplot(arthritis, aes(y = y, x = age)) +
  geom_boxplot(width = 0.5) +
  geom_jitter(alpha = 0.5)
```

Other descriptive plots by category sample size

```{r}
arthritis %>%
  count(trt, y, baseline) %>%
  ggplot(aes(x = baseline, y = y, color = baseline)) +
  geom_point(aes(group = baseline, size = n)) +
  facet_wrap(~trt, scales = "free_x",
             labeller = labeller(trt = label_both)) +
  scale_size(breaks = c(1, 2, 4, 6, 8))

arthritis %>%
  count(trt, y) %>%
  ggplot(aes(x = trt, y = y, color = trt)) +
  geom_point(aes(group = trt, size = n)) +
  scale_size(breaks = c(1, 5, 10, 20, 30))
```

#Ordinal regression models

Checking proportional odds assumption (first way)

```{r}
library(pomcheckr)
pomcheck(y ~ sex + age + trt, arthritis)
```

##Cumulative logit model and probit model with the polr function

```{r}
polr_log <- polr(y ~ ., data = arthritis, Hess = T, model = T, 
                 method = "logistic")
summary(polr_log)
confint(polr_log)

polr_pro <- polr(y ~ ., data = arthritis, Hess = T, model = T, 
                 method = "probit")
summary(polr_pro)
confint(polr_pro)
```

##Cumulative logit model and adjacent-categories logit model with the vglm function

```{r}
cumulative <- vglm(y ~ ., data = arthritis, 
                   family = cumulative(link = "logitlink", 
                                       parallel = T, 
                                       reverse = T))
summary(cumulative)
confint(cumulative)

acat <- vglm(y ~ ., data = arthritis, 
             family = acat(link = "loglink", 
                               parallel = T,
                               reverse = F))
summary(acat)
confint(acat)

#Also continuation-ratio and stopping-ratio logit models
cratio <- vglm(y ~ ., data = arthritis, 
               family = cratio(link = "logitlink", 
                               parallel = T,
                               reverse = F))
summary(cratio)

sratio <- vglm(y ~ ., data = arthritis, 
               family = sratio(link = "logitlink", 
                               parallel = T,
                               reverse = F))
summary(sratio)
```

Goodness-of-Fit and model selection indices for polr and vglm class objects. With the `LR.test`, we can also test the proportional odds assumption.

```{r}
#Normality
library(dgof)
residuals <- residuals(cumulative)
shapiro.test(residuals)
qqnorm(arthritis$age)
qqline(arthritis$age)

#Proportional odds assumption
LR.test(cumulative)

#Goodness-of-Fit
library(gofcat)
lipsitz(polr_log)
brant.test(polr_log)

lipsitz(polr_pro)
brant.test(polr_pro)

lipsitz(cumulative)
LR.test(cumulative)

lipsitz(acat)
LR.test(acat)

#Model selection
AIC(polr_log)
BIC(polr_log)

AIC(polr_pro)
BIC(polr_pro)

AIC(cumulative)
BIC(cumulative)

AIC(acat)
BIC(acat)

#The hosmerlem function works for all three model class objects
hosmerlem(polr_log)
hosmerlem(polr_pro)
hosmerlem(cumulative)
hosmerlem(acat)
hosmerlem(brm)
```

##Cumulative logit model with Bayesian estimation with the brm function

```{r}
options(digits = 10)
fam <- brmsfamily(family = "cumulative", link = "logit") #Define the brms family function
brm <- brm(y ~ ., data = arthritis, family = fam)
summary(brm)
```

Goodness-of-fit and model selection indicies for brm class objects

```{r}
library(loo)
loo(brm) #This function only works for brm class objects
loo::waic(brm)
```

#Predictive performance measures

```{r}
pred <- predict(polr_log, arthritis, type = "probs") #Calculate category prediction probabilities
predict_pl <- max.col(pred) #Select the category with maximum probability

pred2 <- predict(polr_pro, arthritis, type = "probs")
predict_pp <- max.col(pred2)

pred3 <- predict(cumulative, arthritis, type = "response")
predict_cum <- max.col(pred3)

pred4 <- predict(acat, arthritis, type = "response")
predict_ac <- max.col(pred4)

pred5 <- predict(brm, arthritis, type = "probs")
predict_brm <- max.col(pred5)

#library(gmodels)
#cross_pl <- CrossTable(arthritis$y, predict_cat)

#Mutate the prediction to ordered factor
predict_pl <- ordered(predict_pl, levels = ordre_y)
predict_pp <- ordered(predict_pp, levels = ordre_y)
predict_cum <- ordered(predict_cum, levels = ordre_y)
predict_ac <- ordered(predict_ac, levels = ordre_y)
predict_brm <- ordered(predict_brm, levels = ordre_y)

#Confusion matrix
library(caret)
conf_pl <- confusionMatrix(predict_pl, arthritis$y)
conf_pp <- confusionMatrix(predict_pp, arthritis$y)
conf_cum <- confusionMatrix(predict_cum, arthritis$y)
conf_ac <- confusionMatrix(predict_ac, arthritis$y)
conf_brm <- confusionMatrix(predict_brm, arthritis$y)
```

#Intercepts and coefficients estimated parameters

```{r}
#Model coefficients
edf <- data.frame(
  Models = rep(c("polr_log", "polr_pro", "cumulative", "acat", "brm"),
              each = length(polr_log$coefficients)),
  variables = names(polr_log$coefficients),
  coeficients = c(polr_log$coefficients,
                  polr_pro$coefficients,
                  coef(cumulative)[5:8],
                  coef(acat)[5:8],
                  fixef(brm)[5:8,1]),
  lower_CI = c(confint(polr_log)[,1],
               confint(polr_pro)[,1],
               confint(cumulative)[5:8,1],
               confint(acat)[5:8,1],
               fixef(brm)[5:8,3]),
  upper_CI = c(confint(polr_log)[,2],
               confint(polr_pro)[,2],
               confint(cumulative)[5:8,2],
               confint(acat)[5:8,2],
               fixef(brm)[5:8,4])
)

#Intercepts
edf2 <- data.frame(
  Models = rep(c("polr_log", "polr_pro", "cumulative", "acat", "brm"),
              each = 4),
  variables = c("Very poor|Poor", "Poor|Fair", "Fair|Good", "Good|Very good"),
  coefficients = c(c(-1.401, 0.743, 2.585, 4.529),
                  c(-0.848, 0.279, 1.365, 2.486),
                  coef(cumulative)[1:4],
                  coef(acat)[1:4],
                  fixef(brm)[1:4,1]),
  lower_CI = c(c(-2.832, -0.600, 1.212, 3.091),
               c(-1.632, -0.482, 0.593, 1.695),
               confint(cumulative)[1:4,1],
               confint(acat)[1:4,1],
               fixef(brm)[1:4,3]),
  upper_CI = c(c(0.029, 2.086, 3.957, 5.967),
               c(-0.064, 1.039, 2.136, 3.277),
               confint(cumulative)[1:4,2],
               confint(acat)[1:4,2],
               fixef(brm)[1:4,4])
)

edf2$variables <- factor(edf2$variables, 
    levels = c("Very poor|Poor", "Poor|Fair", "Fair|Good", "Good|Very good"))
```

Representation of IC 95%

```{r}
#Coefficients
ggplot(edf, aes(x = variables, y = coeficients, color = Models)) +
  geom_point(position = position_dodge(width = 0.8), size = 3) +
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), 
                position = position_dodge(width = 0.8), width = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  labs(x = "Variables",
       y = "Coefficients") +
  theme_minimal() +
  theme(text = element_text(size = 13))

#Intercepts
ggplot(edf2, aes(x = variables, y = coefficients, color = Models)) +
  geom_point(position = position_dodge(width = 0.8), size = 3) +
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), 
                position = position_dodge(width = 0.8), width = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  labs(x = "Variables",
       y = "Coefficients") +
  theme_minimal() +
  theme(text = element_text(size = 13))

#Alternative to make a plot of coefficients IC for one model

#names(edf) <- c("Estimate","lb","ub")
#require(ggplot2)
#p <- ggplot(aes(y=Estimate,ymin=lb,ymax=ub,x=row.names(edf)),
#data=edf) + geom_pointrange()
#p + coord_flip() + xlab("Predictor") +
#geom_hline(yintercept=0, col=gray(0.75))
```

#Plot of the results of the dataset analysis

```{r, fig.height=5, fig.width=4.5}
levels(arthritis$trt) <- c("No", "Yes")

ggplot(arthritis, aes(x = trt, fill = y)) +
  geom_bar(position = "fill", stat = "count") +
  labs(title = "Self-assessment score of rheumatoid arthritis",
       x = "Treatment",
       y = "Probability") +
  scale_fill_manual(values = c("darkorchid4", "darkorchid1","lightpink", "cadetblue2", "darkcyan"), 
                    name = NULL) +
  theme_minimal() +
  theme(text = element_text(size = 14),
        title = element_text(size = 10),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 13),
    axis.text.y = element_text(size = 10),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.position = "top")

arthritis$baseline <- as.factor(arthritis$baseline)

ggplot(arthritis, aes(x = baseline, fill = y)) +
  geom_bar(position = "fill", stat = "count") +
  labs(title = "Self-assessment score of rheumatoid arthritis",
       x = "Baseline",
       y = "Probability") +
  scale_fill_manual(values = c("darkorchid4", "darkorchid1","lightpink", "cadetblue2", "darkcyan"), 
                    name = NULL) +
  theme_minimal() +
  theme(text = element_text(size = 14),
        title = element_text(size = 10),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 13),
    axis.text.y = element_text(size = 10))
```