---
title: "Systematic Review of R packages for Ordinal Response Regression Models"
author: "Sergi Pujol Rigol"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Search for ordinal regression fitting packages in R CRAN repository 

**Function findFn**

```{r}
library(sos)
library(dplyr)
library(utils)

### Selection 1
selection_1 <- findFn("{ordinal regression}", sortby = NULL, verbose = 1)
selection_1 <- unique(selection_1$Package)

### Selection 2

# Generate a list of relevant keywords
keywords <- c("ordinal", "Ordinal", "ordered", "Ordered", "regression", 
              "Regression", "model", "Model", "logit", "Logit", "logistic", 
              "Logistic", "probit", "Probit", "categorical", "Categorical",
              "response", "Response", "GLMM", "glmm", "GLM", "glm")

# Pairs to avoid
avoid_pairs <- list(
  c("ordinal", "Ordinal"),
  c("ordered", "Ordered"),
  c("regression", "Regression"),
  c("model", "Model"),
  c("logit", "Logit"),
  c("logistic", "Logistic"),
  c("probit", "Probit"),
  c("categorical", "Categorical"),
  c("response", "Response"),
  c("GLMM", "glmm"),
  c("GLM", "glm"),
  c("GLM", "GLMM"),
  c("GLM", "glmm"),
  c("GLMM", "glm"),
  c("glmm", "glm"),
  c("ordinal", "ordered"),
  c("Ordered", "Ordinal"),
  c("Ordinal", "ordered"),
  c("Ordered", "ordinal")
)

# Perform multiple findFn searches, combining three keywords at a time to 
# identify packages that included at least three of them in their descriptions. 
comb <- combn(keywords, 3, simplify = FALSE)
include_keywords <- c("Ordinal", "ordinal", "Ordered", "ordered")
comb_ord <- Filter(function(combinations) {
  any(include_keywords %in% combinations)
}, comb)

valid_comb <- Filter(function(combinations) {
  !any(sapply(avoid_pairs, function(pair) all(pair %in% combinations)))
}, comb_ord)

comb_def <- sapply(valid_comb, function(pair) paste(pair, collapse = " "))

result <- lapply(comb_def, function(combinations) {
  a <- findFn(combinations, sortby = NULL, verbose = 1)
  a <- unique(a$Package)
})

# Filter packages appearing in at least seven different keyword combinations, 
# reducing false positives while maintaining relevant results.
packages <- unlist(result)
count <- table(packages)
packages_list <- names(count[count >= 7])
packages_list

# Generate a dataframe with packages' title and description.
best_results <- data.frame()
for (i in 1:length(packages_list)) {
  url <- paste0("https://cran.r-project.org/web/packages/", 
                packages_list[i], "/index.html")
  
  web_content <- rvest::read_html(url)
  
  package_title <- web_content %>%
    rvest::html_element("h2") %>%
    rvest::html_text(trim = TRUE)

  description <- web_content %>%
    rvest::html_element("p") %>%
    rvest::html_text(trim = TRUE) %>%
    paste(collapse = " ")
  pack <- c(package_title, description)
  print(pack)
    
  best_results <- rbind(best_results, pack)
}

colnames(best_results) <- c("Title", "Description")

best_results <- as.data.frame(lapply(best_results, function(col) {
  if (is.character(col)) {
    gsub("\t", " ", col)
  } else {
    col
  }
}), stringsAsFactors = FALSE)

# Exclude packages whose descriptions did not contain at least one of the 
# following terms: “ordered”, “ordinal”, “regression” or “model”.
best_results_ord <- best_results[grepl("ordered|ordinal|Ordered|Ordinal|
                                       regression|Regression|Model|model", 
                                       best_results[,2], ignore.case = TRUE), ]

write.xlsx(best_results_ord, "/best_results_ord.xlsx") #Revise the list manually

selection_2 <- c(sub(":.*", "", best_results_ord$Title))
```

**Manual search on CRAN package list**

```{r}
selection_3 <- c("BayesOrdDesign","bgms","BinOrdNonNor","bqror","clordr","crov",
                 "CUB","discnorm", "DiscreteGapStatistic", "drord","glmnetcr",
                 "glmpathcr","lcpm","lmmot","Loe","lordif","MABOUST","mgee2",
                 "multgee","multiOrd","multOrdRS","mvord","ohenery","ohoegdm",
                 "opa", "OPSR", "ordCD","ordDisp","ordgam","ordinal",
                 "ordinalbayes","ordinalClust","ordinalCont","ordinalForest",
                 "ordinalgmifs","ordinalLBM","ordinalNet","ordinalpattern",
                 "ordinalRR","OrdNor","ordPens","ORION","ormPlot","otsfeatures",
                 "Passo","PoisBinOrd","PoisBinOrdNonNor","PoisBinOrdNor",
                 "RankAggregator","rankinPlot","Raters","Remiod","rpartScore",
                 "sapevom","Sure","toOrdinal","brms","rms","MASS","VGAM","MXM",
                 "gnlm","lme4","GLMMadaptative","psychotools","jrt","Hmisc",
                 "Splus2R")
```

**Search on R-Universe with "ordinal regression"**

```{r}
selection_4 <- c("VGAM", "MXM","rstanarm","Hmisc","rcompanion","ordinalbayes",
                 "ordinalNet","BayesOrdDesign","ormPlot","wisclabmisc",
                 "monoreg","RcmdrPlugin.NMBU","ordinal","mvord","ordPens",
                 "ordinalgmifs","gnlm","ordinalCont","lordif","ordinalForest",
                 "glmmLasso","ordDisp","reportRmd","MPDiR")
```

Remove repeated packages

```{r}
ds <- c(selection_1, selection_2, selection_3, selection_4)
ds_unique <- unique(ds)
ds_repeated <- duplicated(ds)
ds_repeated <- unique(ds[ds_repeated])
print(ds[ds_repeated])
```

#Downloads count per year and plot

Cumulative number of downloads of all the selected packages from 25 January 2015 to 25 January 2025.

```{r}
library(cranlogs)
library(lubridate)
library(dplyr)
library(ggplot2)

packages <- c("brms", "VGAM", "MachineShop", "rms", "multgee", "tram", 
              "glmmLasso", "ordinalgmifs", "MASS", "ordinal", 
              "BayesOrdDesign", "arm", "ordinalbayes", "ordinalNet", 
              "mvord", "ordinalCont", "gnlm", "oglmx","MCMCglmm", "rstanarm",
              "bqror", "drord", "glmnetcr", "glmpathcr",
              "hopit", "monoreg", "MultOrdRS", "ordgam", "boral", "GLMcat",
              "MNP", "serp", "tramME", "vcrpart", "Rchoice", "geepack",
              "apollo", "plsRglm", "GMMBoost", "crov", "lcpm", "BGLR", "jmv",
              "mixcat", "repolr", "R2MLwiN", "lcmm", "micsr")

df_downloads <- data.frame()
  
for(package in packages){
  downloads <- cran_downloads(package, from = "2015-01-25", to = "2025-01-25")
  df_downloads <- rbind(df_downloads, downloads)
}

df_downloads$year_month <- ym(substr(df_downloads$date, 1, 7))

df_downloads_grouped <- df_downloads %>%
  group_by(year_month, package) %>%
  summarise(count = sum(count))

df_downloads_grouped$ac_count <- ave(df_downloads_grouped$count, 
                                     df_downloads_grouped$package, FUN = cumsum)
df_downloads_grouped$ac_count <- df_downloads_grouped$ac_count/1000

#Group by magnitude of cumulative number of downloads
df_downloads_1 <- df_downloads_grouped %>%
  filter(package %in% c("MASS"))

df_downloads_2 <- df_downloads_grouped %>%
  filter(package %in% c("MCMCglmm","arm","VGAM","rms","ordinal","brms",
                        "geepack","rstanarm","jmv"))

df_downloads_3 <- df_downloads_grouped %>%
  filter(package %in% c("tram","lcmm", "MNP","multgee","BGLR",
                        "apollo", "boral", "Rchoice","ordinalCont", "plsRglm",
                        "glmmLasso"))

df_downloads_4 <- df_downloads_grouped %>%
  filter(package %in% c("ordinalNet","bqror","GMMBoost", "glmnetcr",
                        "MachineShop","R2MLwiN","oglmx","mvord"))

df_downloads_5 <- df_downloads_grouped %>%
  filter(package %in% c("vcrpart","glmpathcr","repolr", "gnlm", 
                        "ordinalgmifs","mixcat","monoreg","tramME","serp",
                        "crov","hopit"))

df_downloads_6 <- df_downloads_grouped %>%
  filter(package %in% c("lcpm","GLMcat","drord","BayesOrdDesign","MultOrdRS",
                        "ordgam","ordinalbayes","micsr"))
```

Plots of the cumulative downloads

```{r, fig.height=5, fig.width=5.5}
ndownload_packages <- c("MASS","arm","rms","VGAM","ordinal","brms","geepack",
                        "rstanarm","MCMCglmm","jmv","tram","lcmm", "MNP",
                        "multgee","BGLR","apollo", "boral", "Rchoice",
                        "ordinalCont", "plsRglm","glmmLasso","ordinalNet",
                        "bqror","GMMBoost", "glmnetcr",
                        "MachineShop","R2MLwiN","oglmx","mvord","vcrpart",
                        "glmpathcr","OrdFacReg","repolr", "gnlm", 
                        "ordinalgmifs","mixcat","monoreg","tramME","serp",
                        "crov","hopit","lcpm","GLMcat","drord","BayesOrdDesign",
                        "MultOrdRS","ordgam","ordinalbayes","micsr")

custom_colors <- c("red2","palegreen4","mediumorchid1","tan","gold",
                     "mediumpurple","palegreen2","hotpink","sienna",
                     "steelblue1","orange","blue4", "lightpink")

par(mfrow = c(3,2))

plot1 <- ggplot(df_downloads_1, 
                aes(x = year_month, y = ac_count, 
                    color = factor(package, levels = ndownload_packages))) +
  geom_path(linewidth = 1) +
  labs(x = "Year",
       y = expression(paste("Number of downloads (x ", 10^3, ")"))) +
  theme_minimal() +
  xlim(as.Date("2015-01-25"),as.Date("2025-01-25")) +
  scale_colour_manual(values = custom_colors) +
  theme(text = element_text(size = 16), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), 
        plot.title = element_text(size = 20), 
        legend.text = element_text(size = 16),
        legend.position = "top",
        legend.title = element_blank())

plot2 <- ggplot(df_downloads_2, 
                aes(x = year_month, y = ac_count, 
                    color = factor(package, levels = ndownload_packages))) +
  geom_path(linewidth = 1) +
  labs(x = "Year",
       y = expression(paste("Number of downloads (x ", 10^3, ")"))) +
  theme_minimal() +
  xlim(as.Date("2015-01-25"),as.Date("2025-01-25")) +
  scale_colour_manual(values = custom_colors) +
  theme(text = element_text(size = 16), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), 
        plot.title = element_text(size = 20), 
        legend.text = element_text(size = 16),
        legend.position = "top",
        legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 3))

plot3 <- ggplot(df_downloads_3, 
                aes(x = year_month, y = ac_count, 
                    color = factor(package, levels = ndownload_packages))) +
  geom_path(linewidth = 1) +
  labs(x = "Year",
       y = expression(paste("Number of downloads (x ", 10^3, ")"))) +
  theme_minimal() +
  xlim(as.Date("2015-01-25"),as.Date("2025-01-25")) +
  scale_colour_manual(values = custom_colors) +
  theme(text = element_text(size = 16), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), 
        plot.title = element_text(size = 20), 
        legend.text = element_text(size = 14),
        legend.position = "top",
        legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 3))

plot4 <- ggplot(df_downloads_4, 
                aes(x = year_month, y = ac_count, 
                    color = factor(package, levels = ndownload_packages))) +
  geom_path(linewidth = 1) +
  labs(x = "Year",
       y = expression(paste("Number of downloads (x ", 10^3, ")"))) +
  theme_minimal() +
  xlim(as.Date("2015-01-25"),as.Date("2025-01-25")) +
  scale_colour_manual(values = custom_colors) +
  theme(text = element_text(size = 16), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), 
        plot.title = element_text(size = 20), 
        legend.text = element_text(size = 14),
        legend.position = "top",
        legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 3))

plot5 <- ggplot(df_downloads_5, 
                aes(x = year_month, y = ac_count, 
                    color = factor(package, levels = ndownload_packages))) +
  geom_path(linewidth = 1) +
  labs(x = "Year",
       y = expression(paste("Number of downloads (x ", 10^3, ")"))) +
  theme_minimal() +
  xlim(as.Date("2015-01-25"),as.Date("2025-01-25")) +
  scale_colour_manual(values = custom_colors) +
  theme(text = element_text(size = 16), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), 
        plot.title = element_text(size = 20), 
        legend.text = element_text(size = 14),
        legend.position = "top",
        legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 3))

plot6 <- ggplot(df_downloads_6, 
                aes(x = year_month, y = ac_count, 
                    color = factor(package, levels = ndownload_packages))) +
  geom_path(linewidth = 1) +
  labs(x = "Year",
       y = expression(paste("Number of downloads (x ", 10^3, ")"))) +
  theme_minimal() +
  xlim(as.Date("2015-01-25"),as.Date("2025-01-25")) +
  scale_colour_manual(values = custom_colors) +
  theme(text = element_text(size = 16), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), 
        plot.title = element_text(size = 20), 
        legend.text = element_text(size = 14),
        legend.position = "top",
        legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 3))

wd <- ""
ggsave(paste0(wd, "down1.png"), plot = plot1, width = 8, height = 6)
ggsave(paste0(wd, "down2.png"), plot = plot2, width = 8, height = 6)
ggsave(paste0(wd, "down3.png"), plot = plot3, width = 8, height = 6)
ggsave(paste0(wd, "down4.png"), plot = plot4, width = 8, height = 6)
ggsave(paste0(wd, "down5.png"), plot = plot5, width = 8, height = 6)
ggsave(paste0(wd, "down6.png"), plot = plot6, width = 8, height = 6)
```

#Creation and updates plot

From a created dataset with the years of creation and updates of the different packages. Only ordinal regression specific packages were included.

```{r}
library(egg)

updates <- data.frame()
count <- 0

for(year in 2009:2025){
  for(i in 1:ncol(Update_year2)){
    for(j in 1:nrow(na.omit(Update_year2[,i]))){
      if(Update_year2[j,i] == year){
        count <- count + 1
      } else {
        count <- count
      }
    }
  }
  df <- c(year, count)
  updates <- rbind(updates, df)
  count <- 0
}

colnames(updates) <- c("Year", "Count")
updates$Year <- as.factor(updates$Year)

ggplot(updates, aes(x = as.numeric(Year), y = Count)) +
  geom_bar(stat = "identity", fill = "orange", alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE, color = "black") +
  labs(title = "Updates per year",
       x = "Year",
       y = "Count") +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 12))
```