---
title: "Ordinal regression analysis"
author: "Sergi Pujol Rigol"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readxl)
library(ggplot2)

library(brms)
library(VGAM)
library(ordinal)
library(MASS)
library(ordinalgmifs)
library(MuMIn)
library(car)
```

#Import and process the injury severity dataset

```{r}
inj_severity <- read_excel("Injury severity.xlsx")

inj_severity <- inj_severity %>%
  dplyr::select(c(id, Categoria, Posicio_camp, Youth_Professional, Temporada, 
           Edat_Final_Temporada, Pes, Talla, Lesio,
           days2recov, `Total_exposure_Hores temporada`)) %>%
  mutate(across(c(Categoria, Posicio_camp, Youth_Professional, Temporada, 
                  Lesio), as.factor)) %>%
  mutate(across(c(Pes, Talla), as.numeric)) %>%
  mutate(id = as.character(id)) %>% #id as character
  rename(Exp = `Total_exposure_Hores temporada`) %>%
  rename(Pos = Posicio_camp) %>%
  rename(Age = Edat_Final_Temporada) %>%
  rename(Y_P = Youth_Professional) %>%
  rename(Ssn = Temporada) %>%
  rename(Hgt = Talla) %>%
  rename(Wgt = Pes)
```

Creation of variable BMI

```{r}
inj_severity$BMI <- inj_severity$Wgt/(inj_severity$Hgt/100)^2
```

Division of the variable `days2recov` in ordered intervals

```{r}
intervals <- c(0,1,8,29,Inf)
inj_severity$Sev <- cut(inj_severity$days2recov, breaks = intervals, 
                        labels = c("None", "Minor", "Moderate", "Serious"),
                        right = FALSE, ordered_result = TRUE)

#Second option, collapsing the Minor and Moderate categories
intervals <- c(0,1,29,Inf)
inj_severity$Sev <- cut(inj_severity$days2recov, breaks = intervals, 
                        labels = c("None", "Minor-Moderate", "Serious"),
                        right = FALSE, ordered_result = TRUE)

summary(inj_severity)
```

Divide the dataset into two subsets for senior (sev_adult) and younger (sev_child) players

```{r}
sev_adult <- subset(inj_severity, Categoria == "Sènior") %>%
  dplyr::select(id, Pos, Y_P, Ssn, Age, Wgt, Hgt, Exp, BMI, Sev)

#All youth/amateur
sev_child <- subset(inj_severity, Categoria != "Sènior") %>%
  dplyr::select(id, Pos, Ssn, Age, Wgt, Hgt, Exp, BMI, Sev)

sev_child <- subset(sev_child, is.na(sev_child$BMI) == FALSE)
sum(is.na(sev_child))
```

#Data description

```{r}
str(sev_adult)
summary(sev_adult)

str(sev_child)
summary(sev_child)
```

Correlations between pairs of variables with the function ggpairs

```{r}
library(GGally)
#Subset senior
ggpairs1 <- GGally::ggpairs(sev_adult[c("Wgt", "Hgt", "BMI")])
ggpairs1 + theme(
  text = element_text(size = 20)
)

ggpairs2 <- GGally::ggpairs(sev_adult[c("Y_P", "Age", "Exp", "BMI")])
ggpairs2 + theme(
  text = element_text(size = 20),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10)
)

#Subset Youth
ggpairs3 <- GGally::ggpairs(sev_child[c("Wgt", "Hgt", "BMI")])
ggpairs3 + theme(
  text = element_text(size = 20)
)
ggpairs4 <- GGally::ggpairs(sev_child[c("Age", "Exp", "BMI")])
ggpairs4 + theme(
  text = element_text(size = 20)
)
```

Choose between BMI, weight and height for both subsets

```{r}
sen_weight <- clmm(Sev ~ Pos + Age + Y_P + Exp + Wgt + Ssn + (1|id), 
       data = sev_adult, Hess = T, model = T, link = "logit", na.action = "na.fail")

sen_height <- clmm(Sev ~ Pos + Age + Y_P + Exp + Hgt + Ssn + (1|id), 
       data = sev_adult, Hess = T, model = T, link = "logit", na.action = "na.fail")

sen_BMI <- clmm(Sev ~ Pos + Age + Y_P + Exp + BMI + Ssn + (1|id), 
       data = sev_adult, Hess = T, model = T, link = "logit", na.action = "na.fail")

you_weight <- clmm(Sev ~ Pos + Age + Exp + Wgt + Ssn + (1|id), 
       data = sev_child, Hess = T, model = T, link = "logit", na.action = "na.fail")

you_height <- clmm(Sev ~ Pos + Age + Exp + Hgt + Ssn + (1|id), 
       data = sev_child, Hess = T, model = T, link = "logit", na.action = "na.fail")

you_BMI <- clmm(Sev ~ Pos + Age + Exp + BMI + Ssn + (1|id), 
       data = sev_child, Hess = T, model = T, link = "logit", na.action = "na.fail")

anova(sen_weight, sen_height)
anova(sen_height,sen_BMI)
anova(sen_weight,sen_BMI)

anova(you_weight, you_height)
anova(you_height, you_BMI)
anova(you_weight, you_BMI)
```

Backward stepwise regression with function `drop1` and model selection with function `dredge`, includin Ssn as fixed effect.

```{r}
mod_adult <- clmm(Sev ~ Pos + Age + Y_P + Exp + Wgt + Ssn + (1|id), 
       data = sev_adult, Hess = T, model = T, link = "logit", na.action = "na.fail")

mod_child <- clmm(Sev ~ Pos + Age + Exp + BMI + Ssn + (1|id), 
       data = sev_child, Hess = T, model = T, link = "logit", na.action = "na.fail")
```

```{r}
dredge_adult <- dredge(mod_adult)
dredge_child <- dredge(mod_child)

drop1(mod_adult, trace = T, test = "Chisq")
drop1(mod_child, trace = T, test = "Chisq")
```

ANOVA to compare Season as fixed effect and random effect.

```{r}
mod1_senior <- clmm(Sev ~ Pos + Age + Y_P + BMI + Ssn + (1|id), 
       data = sev_adult, Hess = T, model = T, link = "logit", na.action = "na.fail")

mod2_senior <- clmm(Sev ~ Pos + Age + Y_P + BMI + (1|Ssn) + (1|id), 
       data = sev_adult, Hess = T, model = T, link = "logit", na.action = "na.fail")

anova(mod2_senior, mod1_senior)

mod1_youth <- clmm(Sev ~ Pos + Age + BMI + Ssn + (1|id), 
       data = sev_child, Hess = T, model = T, link = "logit", na.action = "na.fail")

mod2_youth <- clmm(Sev ~ Pos + Age + BMI + (1|Ssn) + (1|id), 
       data = sev_child, Hess = T, model = T, link = "logit", na.action = "na.fail")

anova(mod2_youth, mod1_youth)
```

#Fit cumulative logit mixed-effects models

With the clmm function from the ordinal package for both subsets.

```{r}
#Seniors' subset
mod1_senior <- clmm(Sev ~ Pos + Age + Y_P + BMI + Ssn + (1|id), 
       data = sev_adult, Hess = T, model = T, link = "logit", na.action = "na.fail")
summary(mod1_senior)
confint(mod1_senior)

#Youth's subset
mod2_youth <- clmm(Sev ~ Pos + Age + BMI + (1|Ssn) + (1|id), 
       data = sev_child, Hess = T, model = T, link = "logit", na.action = "na.fail")
summary(mod2_youth)
confint(mod2_youth)
```

Proportional odds assumption test

```{r}
mod3_youth <- clm(Sev ~ Pos + Age + BMI, 
       data = sev_child, Hess = T, model = T, link = "logit", na.action = "na.fail")
library(gofcat)
brant.test(mod3_youth)

require(rms)
y <- as.factor(inj_severity$Sev)
Y <- as.numeric(y) - 1
ncut <- length(unique(Y)) - 1
p <- 5
Coef <- matrix(NA, ncol=p, nrow=ncut,
               dimnames=list(paste('>=', levels(y)[-1],sep=''),
                 NULL))
for(k in 1:ncut) {
  f <- lrm(Y >= k ~ Pos + Y_P + Ssn + Age + BMI, data=inj_severity)
  Coef[k,] <- coef(f)[-1]
}
colnames(Coef) <- names(coef(f))[-1]
round(Coef, 3)
```

Plots of the intercepts and coefficients estimations and IC 95%

```{r}
#IC 95% estimations intercepts and coefficients for seniors' model
edf1 <- data.frame(
  Estimations = c(rep("Intercepts", 3),rep("Coefficients",9)),
  variables = names(mod1_senior$coefficients),
  coefficients = mod1_senior$coefficients,
  lower_CI = confint(mod1_senior)[,1],
  upper_CI = confint(mod1_senior)[,2]
)
order <- names(mod1_senior$coefficients)
edf1$variables <- factor(edf1$variables, levels = order)

ggplot(edf1, aes(x = variables, y = coefficients, color = Estimations)) +
  geom_point(position = position_dodge(width = 0.8), size = 3) +
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), 
                position = position_dodge(width = 0.8), width = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_fill_manual(name = NULL) +
  labs(x = "Variables",
       y = "Coefficients") +
  theme_minimal() +
  theme(text = element_text(size = 13),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "top")

#IC 95% estimations intercepts and coefficients for seniors' model

edf2 <- data.frame(
  Estimations = c(rep("Intercepts", 3),rep("Coefficients",5)),
  variables = names(mod2_youth$coefficients),
  coefficients = mod2_youth$coefficients,
  lower_CI = confint(mod2_youth)[,1],
  upper_CI = confint(mod2_youth)[,2]
)
order2 <- names(mod2_youth$coefficients)
edf2$variables <- factor(edf2$variables, levels = order2)

ggplot(edf2, aes(x = variables, y = coefficients, color = Estimations)) +
  geom_point(position = position_dodge(width = 0.8), size = 3) +
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI), 
                position = position_dodge(width = 0.8), width = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_fill_manual(name = NULL) +
  labs(x = "Variables",
       y = "Coefficients") +
  theme_minimal() +
  theme(text = element_text(size = 13),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "top")
```

#Plot of the results of the dataset analysis

```{r, fig.height = 5, fig.width=5}
#Severity vs category (Y_P) in seniors' subset
levels(sev_adult$Y_P) <- c("Amateur", "Professional")

ggplot(sev_adult, aes(x = Y_P, fill = Sev)) +
  geom_bar(position = "fill", stat = "count") +
  labs(title = "Gráfico de Barras Apiladas",
       x = "Player level",
       y = "Probability") +
  scale_fill_manual(values = c("blue4", "cornflowerblue", "coral", "brown3"), 
                    name = "Injury severity") +
  theme_minimal() +
  theme(text = element_text(size = 14),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 13),
    axis.text.y = element_text(size = 10),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12))

#Severity vs season in seniors' subset
levels(sev_adult$Ssn) <- c("2017-18", "2018-19", "2019-20", "2020-21")

ggplot(sev_adult, aes(x = Ssn, fill = Sev)) +
  geom_bar(position = "fill", stat = "count") +
  labs(title = "Injury severity",
       x = "Season",
       y = "Proportion") +
  scale_fill_manual(values = c("blue4", "cornflowerblue", "coral", "brown3"), name = NULL) +
  theme_minimal() +
  theme(text = element_text(size = 14),
        plot.title = element_text(size = 14),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 13, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10),
    legend.text = element_text(size = 12),
    legend.position = "top")
```

```{r, fig.height = 6, fig.width=5}
#Severity vs Position in youth's subset
levels(sev_child$Pos) <- c("Winger", "Forward", "Goalkeeper", "Defender")

ggplot(sev_child, aes(x = Pos, fill = Sev)) +
  geom_bar(position = "fill", stat = "count") +
  labs(title = "Injury severity",
       x = "Player position",
       y = "Probability") +
  scale_fill_manual(values = c("blue4", "cornflowerblue", "coral", "brown3"), 
                    name = NULL) +
  theme_minimal() +
  theme(text = element_text(size = 14),
        plot.title = element_text(size = 14),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 13, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.position = "top")

#Severity vs BMI in youth's subset
ggplot(sev_child, aes(x = BMI, fill = Sev)) +
  geom_density(position = "fill") +
  labs(x = "Body Mass Index",
       y = "Probability") +
  scale_fill_manual(values = c("blue4", "cornflowerblue", "coral", "brown3"), 
                    name = "Injury severity") +
  theme_minimal() +
  theme(text = element_text(size = 14),
        plot.title = element_text(size = 14),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 13),
    axis.text.y = element_text(size = 10),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)) +
  scale_x_continuous(breaks = seq(14, 30, by = 2))
```