---
title: "TFM"
author: "Sergi Pujol Rigol"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Search for ordinal regression fitting packages in CRAN 

**Function findFn**

```{r}
library(sos)

selection_1 <- findFn("{ordinal regression}", sortby = NULL, verbose = 1)
selection_2 <- findFn("ordered response factor", sortby = NULL, verbose = 1)

selection_1 <- unique(selection_1$Package)
selection_2 <- unique(selection_2$Package)
```

**Manual search on CRAN package list**

```{r}
selection_3 <- c("BayesOrdDesign","Bgms","BinOrdNonNor","Bqror","Clordr","Crov",
                 "CUB","discnorm","Drord","glmnetcr","glmpathcr","Lcpm","Lmmot",
                 "Loe","Lordif","MABOUST","mgee2","Multgee","MultiOrd","MultOrdRS",
                 "Mvord","Ohenery","ohoegdm","Opa","OrdCD","ordDisp","Ordgam",
                 "Ordinal","ordinalbayes","ordinalClust","ordinalCont","ordinalForest",
                 "ordinalgmifs","ordinalLBM","ordinalNet","ordinalpattern",
                 "ordinalRR","OrdNor","ordPens","ORION","ormPlot","otsfeatures",
                 "Passo","PoisBinOrd","PoisBinOrdNonNor","PoisBinOrdNor",
                 "RankAggregator","rankinPlot","Raters","Remiod","rpartScore",
                 "sapevom","Sure","toOrdinal","Brms","Rms","MASS","VGAM","MXM",
                 "Gnlm","lme4","GLMMadaptative","psychotools","Jrt","Hmisc","Splus2R")
```

**Search on R-Universe with "ordinal regression"**

```{r}
selection_4 <- c("VGAM", "MXM","rstanarm","Hmisc","rcompanion","ordinalbayes",
                 "ordinalNet","BayesOrdDesign","ormPlot","wisclabmisc",
                 "monoreg","RcmdrPlugin.NMBU","ordinal","mvord","ordPens",
                 "ordinalgmifs","gnlm","ordinalCont","lordif","ordinalForest",
                 "glmmLasso","ordDisp","reportRmd","MPDiR")
```

Remove repeated packages

```{r}
ds <- c(selection_1, selection_2, selection_3, selection_4)
ds_unique <- unique(ds)
ds_repeated <- duplicated(ds)
ds_repeated <- unique(ds[ds_repeated])
print(ds[ds_repeated])
```

#Other searches with findFn function

These results were not included, but can be included in a future perspective.

```{r}
a <- findFn("{ordered factor}")
b <- findFn("ordered response factor")
c <- findFn("proportional odds model")
d <- findFn("continuation ratio model")
e <- findFn("adjacent categories model")
f <- findFn("partial credit model")
g <- findFn("fit ordinal model")

a <- unique(a$Package)
b <- unique(b$Package)
c <- unique(c$Package)
d <- unique(d$Package)
e <- unique(e$Package)
f <- unique(f$Package)
g <- unique(g$Package)

packages <- c(a,b,c,d,e,f,g)
n_packages <- as.data.frame(table(packages))

findFn("{regression model}", sortby = NULL, verbose = 1)
findFn("continuation-ratio model")
```

#Downloads count per year and plot

Cumulative number of downloads of all the selected packages from 1 November 2013 to 10 November 2023

```{r}
library(cranlogs)
library(lubridate)
library(dplyr)
library(ggplot2)

packages <- c("brms", "VGAM", "MachineShop", "rms", "multgee", "tram", 
              "glmmLasso", "ordinalgmifs", "MASS", 
              "jrt", "ordinal", "BayesOrdDesign", "arm", "MXM", "ordinalbayes",
              "ordinalNet", "mvord", "ordinalCont", "gnlm", "oglmx", "lordif",
              "OrdFacReg", "MCMCglmm", "rstanarm")

df_downloads <- data.frame()
  
for(package in packages){
  downloads <- cran_downloads(package, from = "2013-11-01", to = "2023-11-01")
  df_downloads <- rbind(df_downloads, downloads)
}

df_downloads$year_month <- ym(substr(df_downloads$date, 1, 7))

df_downloads_grouped <- df_downloads %>%
  group_by(year_month, package) %>%
  summarise(count = sum(count))

df_downloads_grouped$ac_count <- ave(df_downloads_grouped$count, df_downloads_grouped$package, FUN = cumsum)
df_downloads_grouped$ac_count <- df_downloads_grouped$ac_count/1000

#Group by magnitude of cumulative number of downloads
df_downloads_1 <- df_downloads_grouped %>%
  filter(package %in% "MASS")

df_downloads_2 <- df_downloads_grouped %>%
  filter(package %in% c("arm","VGAM","rms","ordinal","brms","MCMCglmm","rstanarm"))

df_downloads_3 <- df_downloads_grouped %>%
  filter(package %in% c("ordinalCont", "tram","glmmLasso","MXM","lordif",
                        "multgee"))

df_downloads_4 <- df_downloads_grouped %>%
  filter(!(package %in% c("MASS","arm","VGAM","rms","ordinal","brms","MCMCglmm","rstanarm",
                          "ordinalCont", "tram","glmmLasso","MXM","lordif",
                          "multgee")))
```

Plots of the cumulative downloads

```{r, fig.height=5, fig.width=5.5}
ndownload_packages <- c("MASS","arm","rms","VGAM","ordinal","brms","rstanarm","MCMCglmm",
                        "tram","multgee","MXM","ordinalCont","glmmLasso",
                        "lordif","ordinalNet","MachineShop","oglmx",
                        "OrdFacReg","mvord","gnlm","ordinalgmifs",
                        "jrt","BayesOrdDesign","ordinalbayes")

custom_colors_1 <- c("red2")
custom_colors_2 <- c("gold1","mediumpurple","palegreen2","hotpink",
                   "sienna","steelblue1","orange")
custom_colors_3 <- c("palegreen4","mediumorchid1","tan","tomato3","blue4","gold1")
custom_colors_4 <- c("mediumpurple","palegreen2","hotpink",
                   "sienna","steelblue1","orange","red2","palegreen4",
                   "mediumorchid1","tan")

par(mfrow = c(2,2))

ggplot(df_downloads_1, aes(x = year_month, y = ac_count, color = factor(package, levels = ndownload_packages))) +
  geom_path(linewidth = 1) +
  labs(x = "Year",
       y = expression(paste("no. of downloads (x ", 10^3, ")"))) +
  theme_minimal() +
  xlim(as.Date("2013-11-01"),as.Date("2023-10-31")) +
  scale_colour_manual(values = custom_colors_1) +
  theme(text = element_text(size = 16), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), 
        plot.title = element_text(size = 20), 
        legend.text = element_text(size = 16),
        legend.position = "top",
        legend.title = element_blank())

ggplot(df_downloads_2, aes(x = year_month, y = ac_count, color = factor(package, levels = ndownload_packages))) +
  geom_path(linewidth = 1) +
  labs(x = "Year") +
  theme_minimal() +
  xlim(as.Date("2013-11-01"),as.Date("2023-10-31")) +
  scale_colour_manual(values = custom_colors_2) +
  theme(text = element_text(size = 16), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), 
        plot.title = element_text(size = 20), 
        legend.text = element_text(size = 16),
        legend.position = "top",
        legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 3))

ggplot(df_downloads_3, aes(x = year_month, y = ac_count, color = factor(package, levels = ndownload_packages))) +
  geom_path(linewidth = 1) +
  labs(x = "Year",
       y = expression(paste("no. of downloads (x ", 10^3, ")"))) +
  theme_minimal() +
  xlim(as.Date("2013-11-01"),as.Date("2023-10-31")) +
  scale_colour_manual(values = custom_colors_3) +
  theme(text = element_text(size = 16), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), 
        plot.title = element_text(size = 20), 
        legend.text = element_text(size = 16),
        legend.position = "top",
        legend.title = element_blank())

ggplot(df_downloads_4, aes(x = year_month, y = ac_count, color = factor(package, levels = ndownload_packages))) +
  geom_path(linewidth = 1) +
  labs(x = "Year") +
  theme_minimal() +
  xlim(as.Date("2013-11-01"),as.Date("2023-10-31")) +
  scale_colour_manual(values = custom_colors_4) +
  theme(text = element_text(size = 16), 
        axis.title = element_text(size = 18), 
        axis.text = element_text(size = 14), 
        plot.title = element_text(size = 20), 
        legend.text = element_text(size = 14),
        legend.position = "top",
        legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 4))
```

#Creation and updates plot

From a created dataset with the years of creation and updates of the different packages. Only ordinal regression specific packages were included.

```{r}
library(egg)

updates <- data.frame()
count <- 0

for(year in 2009:2023){
  for(i in 1:ncol(Update_year2)){
    for(j in 1:nrow(na.omit(Update_year2[,i]))){
      if(Update_year2[j,i] == year){
        count <- count + 1
      } else {
        count <- count
      }
    }
  }
  df <- c(year, count)
  updates <- rbind(updates, df)
  count <- 0
}

colnames(updates) <- c("Year", "Count")
updates$Year <- as.factor(updates$Year)

ggplot(updates, aes(x = as.numeric(Year), y = Count)) +
  geom_bar(stat = "identity", fill = "orange", alpha = 0.5) +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE, color = "black") +
  labs(title = "Updates per year",
       x = "Year",
       y = "Count") +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 12))
```